---
title: "Export multiple data frames to Excel workbook"
subtitle: "with openxlsx::write.xlsx() `R` function"
author: |
  | [Ogundepo Ezekiel Adebayo](https://bit.ly/gbganalyst)
  | [I'm on Twitter](https://twitter.com/gbganalyst)
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: simplex
    highlight: espresso
    toc: true
    toc_float: true
    code_download: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`openxlsx` package makes it easier to export multiple tables in `R` to `Excel` workbook with each Excel worksheet corresponds to a data frame in R.


# Package

```{r package, warning=FALSE, message=FALSE}
library(tidyverse)
library(openxlsx)
library(readxl)

```

# Data to export

For example, we want to export `airquality`, `mtcars`, `iris`, and `diamonds` dataset to a different Excel worksheet.

```{r dataset}
dataframe1 <- datasets::airquality
dataframe1 %>% head()

dataframe2 <- datasets::mtcars
dataframe2 %>% head()


dataframe3 <- datasets::iris
dataframe3 %>% head()

dataframe4 <- ggplot2::diamonds
dataframe4 %>% head()
```

Now that we have our data frames in R, we shall name the Excel sheets according to the name of the data frames in R.

* sheet1- airquality
* sheet2- mtcars
* sheet3- iris
* sheet4- diamonds

```{r Export excel, message=FALSE, warning= FALSE}
list_of_datasets <- list("airquality" = dataframe1, "mtcars" = dataframe2, "iris" = dataframe3, "diamonds" = dataframe4)

write.xlsx(list_of_datasets, "Excel workbook.xlsx")
```

We can now look for `Excel workbook.xlsx` file in the working directory.

# Follow up

We can use Microsoft Excel to open **Excel workbook.xlsx** file in the working directory but we deliberately use R to see the worksheets with the help of `readxl::excel_sheets()` function.

```{r excel worksheets}
excel_sheets("Excel workbook.xlsx")
```

In order see what each excel worksheet contains, we use `purrr::map()` function 

```{r}

path <- "Excel workbook.xlsx"

path %>%
        excel_sheets()%>%
        set_names()%>%
       map(read_excel,
           path=path)
        
```

---
If you like this writeup, you can also follow me on [Twitter](https://www.twitter.com/gbganalyst){target="_blank"} and [Linkedin](https://www.linkedin.com/in/ezekiel-ogundepo/){target="_blank"} for more updates in `R` and `Python` for datascience.